---
output: 
  html_document:
    theme: readable
    highlight: textmate
    keep_md: true
---
# Distribution of Application Risk Scores in IBM X-Force Exchange 
### Author:  Brian Matthiesen
### November 2015
_ _ _

<a name="top-page"/>  

## Table of Contents
* [Synopsis](#synopsis)
* [Data Processing](#processing)
* [Results](#results)
    * [Distribution of Risk Scores](#riskdist)
* [Conclusion](#conclusion)

_ _ _ _ _ _ _ _ _ _ _

<a name="synopsis"/>

## Synopsis
In this report we aim to describe the distribution of Internet Application Profile (IAP) Risk Scores within the IBM X-Force Exchange service. Our hypothesis is that the bulk of the applications are clustered within a narrow band of scores at the low end of the range with relatively few high-risk applications at the margin. To investigate this hypothesis, we obtained data on Application (IAP) entries in X-Force Exchange via the public API. From these data we found that the bulk of the application risk scores are in the range of 2 to 4.  

[[Top](#top-page)]

_ _ _ _ _ _ _ _ _ _ _

<a name="processing"/>

## Data Processing
This section describes the source and the preparation of data for the analysis. 
We begin by preparing the R environment by clearing the environment and loading library packages used by this analysis. 
  
```{r lib_load, message=FALSE}
rm(list=ls())
library(ggplot2)
library(RCurl)
library(jsonlite)
library(dplyr)
library(knitr)

```
  
The data for this analysis comes from the IBM X-Force Exchange service. (exchange.xforce.ibmcloud.com). 
The Internet Application Profile (IAP) are available via a RESTful interface and are returned as JSON objects.   
  
```{r data_load, cache=TRUE}
# Get the authorization token from xForce Exchange. Save the token in the curl
# options for use on all xForce API calls.

xfe_auth = fromJSON(getURL("https://xforce-api.mybluemix.net/auth/anonymousToken"))
curl.opts <- list(httpheader = c("Accept-Language" = "en-US",
                                 "Authorization" = paste("Bearer", xfe_auth$token),
                                 "Accept" = "application/json"))

# Call xForce to return the list of all known applications in JSON format.
# JASON key "canonicalNames" holds the list of application names.

appList <- fromJSON(getURL("https://api.xforce.ibmcloud.com/app/",
                           .opts = curl.opts))
```
  
Now that the data are loaded, let's review some of the characteristics of the Application List (IAP). 
The number of Applications found in X-Force is `r length(appList[[1]])`.  We also show the first 10 applications.
  
```{r data_display1}
length(appList[[1]])
head(appList[[1]],10)
```

  
We are interested in the Risk Score for each application. These data are contained in the Application Details. 
We call the X-Force API to retrieve these scores along with additional data such as application category and known risks.  
  
```{r data_display2}
# Create the final receiving data frame to be used in plotting and loop through
# the list of applications known to xForce, retrieving details for each
# application. Set application risk to "5" if it is missing from xForce.

appData <- data.frame()

for (appName in appList$canonicalNames) {
    # Replace spaces in URLs  
    appName <- gsub(" ", "%20", appName)        
    
    # Call xForce to pull the app details
    appDetail <- fromJSON(getURL(paste0("https://xforce-api.mybluemix.net/app/", 
                                        appName), .opts = curl.opts))
    
    # Replace missing risk scores with "5.0"
    if (is.null(appDetail$application$score)) {
        appDetail$application$score <- "5"
    }
    
    # Accumulate the data table for plotting.
    # Note that JSON keys map to R "name" when using fromJSON function.
    appData <- rbind(appData, matrix(list(
                                "Application",
                                appDetail$application$name,
                                appDetail$application$description,
                                names(appDetail$application$categories),
                                names(appDetail$application$actions),
                                names(appDetail$application$rlfs),
                                as.numeric(appDetail$application$score)),
                                nrow = 1, ncol = 7))
}

# Add column names to make it easier to work with.
colnames(appData) <- list("Type", "Name", "Description", "Categories", "Actions", "Rilfs", "Score")

```
  
Now that the Application Detail data are loaded, let's review some of the characteristics.
We first list a few applications and their risk scores, then a full IAP entry showing category and other data.
  
```{r data_display3}
head(appData[, c(2,7)],20)
appData[1,]
```

_ _ _ _ _ _ _ _ _ _ _
<a name="result"/>

## Results

<a name="riskdist"/>

### Distribution of Risk Scores
After collecting all Application (IAP) data, we can summarize the statistics of the Risk Score.
We also plot a histogram of the Risk Score distribution.

```{r risk_dist, fig.height=5, fig.width=10, message=FALSE}
# Summarize the statistics of the risk scores. 
x <- as.vector(appData$Score, mode="numeric")
summary(x)

# Plot the histogram of xForce risk scores. 
h <- hist(x, breaks=10, col="red", xlab="xForce Risk Scores", 
          main="Distribution of Risk Scores for xForce Exchange Applications")

# Add a distribution curve.
xfit <-seq(min(x),max(x),length=40) 
yfit <-dnorm(xfit,mean=mean(x),sd=sd(x)) 
yfit <- yfit*diff(h$mids[1:2])*length(x) 
lines(xfit, yfit, col="blue", lwd=2)
```  
[[Top](#top-page)]   
   
_ _ _ _ _ _ _ _ _ _ _
<a name="recommend"/>

## Recommendations
The data support our hypothesis that the bulk of the applications are clustered within a narrow band of scores at the low end of the range with relatively few high-risk applications at the margin. The data are distributed with a median of `r median(x)` with a few high-risk applications with a maximum of `r max(x)`. 


[[Top](#top-page)]  
